---
title: "hydraR"
format: 
  gfm:
    wrap: none
---

```{r}
#| echo: false
options(hydraR.print.max_depth = 10L, hydraR.print.values = TRUE)
```

Modern data analysis projects need to store and retrieve many configuration settings, such as secret keys and tokens, paths, tuning parameters, database ports, and authentication values.  The `hydraR` package for `R` provides a powerful and flexible composition workflow to manage such configuration settings. This package is written entirely in `R`, but is designed to be (mostly) API-compatible two very popular Python libraries: OmegaConf and Hydra.

OmegaConf is a hierarchical configuration system with interpolation and resolvers ([website](https://omegaconf.readthedocs.io/)).
Hydra is a composition framework built on OmegaConf for defaults-driven configuration and runtime overrides ([website](https://hydra.cc/)).
hydraR is implements a focused Hydra/OmegaConf-style composition workflow for YAML configs in `R`.

## Public API

The user-facing API is intentionally minimal:

```r
compose(config_path, config_name, overrides = ...)
```

## Install

```bash
remotes::install_github("vincentarelbundock/hydraR")
```

## Quick Start

Fixture layout used in these examples:

```text
inst/testdata/conf/
  main.yml
  switch.yml
```

Example:

```{r}
pkgload::load_all()

conf_dir <- system.file("testdata", "conf", package = "hydraR")

cfg <- compose(
  config_path = conf_dir,
  config_name = "switch"
)
cfg
cfg$mode
cfg$run$lr

cfg <- compose(
  config_path = conf_dir,
  config_name = "main",
  overrides = c("run.lr=0.01", "selected_plan=B")
)

cfg
```

## Alternative Configs

```{r}
compose(config_path = conf_dir, config_name = "main")$run$lr
compose(config_path = conf_dir, config_name = "switch")$run$lr
```

## Value Overrides

Set or replace values with `key=value`:

```{r}
cfg <- compose(
  config_path = conf_dir,
  config_name = "main",
  overrides = c("service.retries=7", "selected_plan=B")
)
cfg$service$retries
cfg$plan
```

Use `+key=value` to add a key that must not already exist:

```{r}
cfg <- compose(config_path = conf_dir, config_name = "main", overrides = c("+new_key=abc"))
cfg$new_key
```

Use `++key=value` to upsert (insert or replace):

```{r}
compose(config_path = conf_dir, config_name = "main", overrides = c("++new.branch=7"))$new$branch
```

Use `~key` (or `~key=expected_value`) to delete keys:

```{r}
cfg <- compose(config_path = conf_dir, config_name = "main", overrides = c("~service.retries"))
is.null(cfg$service$retries)
```

`~` can also target nested paths (for example `~service.retries`).

## Config Group Overrides

Config-group selection overrides are supported when your config directory
contains group folders (for example `db/mysql.yaml`, `db/postgres.yaml`)
and defaults entries such as:

```yaml
defaults:
  - db: mysql
  - _self_
```

Then:

- `db=postgres` replaces the selected group option.
- `+db=postgres` inserts a new group selection if absent.
- `++db=postgres` upserts group selection.
- `~db` removes the group defaults entry.

## Environment Variables

Use `${oc.env:VAR_NAME,fallback}` to read from the process environment.

Set variables before composing:

```bash
export APP_USER=alice
export APP_PORT=3310
```

Example config values:

```yaml
db:
  user: ${oc.env:APP_USER,guest}
  password: ${oc.env:APP_PASSWORD,???}
  port: ${oc.decode:${oc.env:APP_PORT,3306}}
```

- `guest` is used if `APP_USER` is unset.
- `???` makes `APP_PASSWORD` required (compose errors if missing).
- `oc.decode` turns `"3306"`/`"3310"` into numeric values.

## Interpolation

`compose()` resolves `${...}` interpolations after composition and overrides.

```{r}
cfg_interp <- compose(
  config_path = conf_dir,
  config_name = "main",
  overrides = c("selected_plan=B")
)

cfg_interp$client$url
cfg_interp$client$description
cfg_interp$plan
```

## Extending Configs

Defaults inside non-primary configs are supported, so a config can extend
another config and override selected fields.

Same-group extension example (`db/mysql_ext.yaml`):

```yaml
defaults:
  - base_mysql_ext
user: omry
password: secret
port: 3307
encoding: utf8
```

Cross-group extension into the current package:

```yaml
defaults:
  - /db_schema/base_mysql@_here_
```

## Development

```bash
make document
make test
```

## Contributing

See `AGENTS.md` for repository contribution guidelines.

Before opening a PR, run:

```bash
make document
make test
make check
```
